# Reproducuble results Peer Assesment project 1

## Adding important packages
```{r}
library(knitr)
opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(ggplot2)
```
## Processing the data
Here we change the format of dates from a factor to POSFIX using lubridate package
```{r}
act <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character","integer"))
act$date <- ymd(act$date)
```

## Calculating number of steps as per date
```{r}

steps <- act %>%
filter(!is.na(steps)) %>%
group_by(date) %>%
summarize(steps = sum(steps)) %>%
print
```

## A histogram showing number of steps on each date ignorging NA values
```{r}
ggplot(steps, aes(x = steps)) + geom_histogram(fill = "blue", binwidth = 1000) + labs(title = "Histogram of steps per day", x = "steps per day", y = "Frequency")
```

##Calculating mean steps and median steps

we see here that these values are different because of missing values.
```{r}
mean_steps <- mean(steps$steps, na.rm = T)
median_steps <- median(steps$steps, na.rm = T)
mean_steps
median_steps
```

## grouping steps according to a fix interval
```{r}

interval <- act %>%
filter(!is.na(steps)) %>%
group_by(interval) %>%
summarize(steps = mean(steps))
```
## Plotting steps for interval
```{r}
ggplot(interval, aes(x = interval, y = steps))+geom_line(color="blue")
interval[which.max(interval$steps),]
```



## Imputing strategy


Strategy for filling in all of the missing values in the dataset:  I am using mean of that day to fill the missing value 

```{r}

sum(is.na(act$steps))
```

##Average daily activity pattern, mutating values for missing values
```{r}

act_full <- act
nas <- is.na(act_full$steps)
avg_interval <- tapply(act_full$steps, act_full$interval, mean, na.rm=T, simplify = T)
act_full$steps[nas] <- avg_interval[as.character(act_full$interval[nas])]
```

```{r}

sum(is.na(act_full$steps))


steps_full <- act_full %>%
filter(!is.na(steps)) %>%
group_by(date) %>%
summarize(steps = sum(steps)) %>%
print

ggplot(steps_full, aes(x=steps)) + geom_histogram(fill="blue", binwidth=1000)+labs(title = "Histogram of steps per day, including missing values", x = "steps per day", y = "Frequency")
mean_steps_full <- mean(steps_full$steps, na.rm = T)
median_steps_full <- median(steps_full$steps, na.rm = T)
mean_steps_full
median_steps_full

```
## adding a column for weektype
```{r}
act_full <- mutate(act_full, weektype = ifelse(weekdays(act_full$date) == "Saturday" | weekdays(act_full$date) == "Sunday", "weekend", "weekday"))
act_full$weektype <- as.factor(act_full$weektype)
head(act_full)

interval_full <- act_full %>%
group_by(interval, weektype) %>%
summarize(steps = mean(steps))
```

## Separte plot for each weektype steps for interval 
```{r}
s <- ggplot(interval_full, aes(x=interval, y = steps, color = weektype)) + geom_line()+facet_wrap(~weektype, ncol = 1 , nrow = 2)
print(s)

```







